A Comparative study of Dask and Spark for neuroimaging pipelines

### 2019-06-28 latex/paper.tex
Worked on Applications section

We used three neuroimaging applications to evaluate the engines in
different conditions. The first two ones, incrementation and histogram, are
simple synthetic benchmarks representing basic map-only and map-reduce applications.
The third one is a more realistic application representative of popular
BIDS applications~\cite{gorgolewski2017bids}.

%% Incrementation
\subsubsection{Incrementation}
We used the simple image incrementation pipeline used
in~\cite{hayot2019performance} (see Algorithm~\ref{alg:incrementation}).
The application reads blocks of the BigBrain image from the shared file
system, increments the intensity value of each voxel by 1 to avoid caching
effects, sleeps for a configurable amount of time to emulate a more complex
processing, repeats this process for a specified amount of iterations, and
finally writes the result as a NifTI image back to the shared file system.
This application allows us to study the behavior of the engines when all
inputs are processed independently, in a map-only way (see
Figure~\ref{fig:tg-inc}). This mimics the behavior of analyzing multiple
independent subjects in parallel.

@@

\begin{figure}[!b]
    \centering
    \includegraphics[width=0.125\textwidth,
    angle=-90]{images/incrementation-task-graph.png}
    \caption{Task graph for Incrementation with 5 iterations, represented
    for 3 BigBrain blocks. Circles represent the incrementation and sleep
    function while rectangles represent image blocks.}\label{fig:tg-inc}
\end{figure}

@@

 Our second application calculates the histogram of the BigBrain image (see
 Algorithm~\ref{alg:histogram}). It reads the image blocks from the shared
 file system, calculates intensity frequencies, aggregates the frequencies
 across the blocks, and finally writes the resulting histogram to the
 shared file system as a single \TG{X}KB file. This application implements
 a typical map-reduce pattern, where the final result is obtained from all
 the input blocks (see Fig.~\ref{fig:tg-histo}). This application requires
 data shuffling thus inter-worker communication. The total amount of shuffled
 data is however limited to \TG{X}MB as it only consists of image
 histograms.

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.16\textwidth, angle=-90]{images/histogram-task-graph.png}
    \caption{Task graph for Histogram}\label{fig:tg-histo}
\end{figure}

@@

We used the \href{https://github.com/BIDS-Apps/example}{BIDS app example}
application available on GitHub. This application operates in a map-reduce
way. The map phase, also called participant analysis, extracts the brain
tissues of a subject's 3D MRI using the popular FMRIB Software Library
(\href{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki}{FSL}), and writes the
resulting image (\TG{X}MB on average) to the shared file system. The reduce
phase, also called group analysis, computes the volume of each brain and
returns the average volume, shuffling a total of \TG{X}GB image data.

@@

We varied four parameters in our experiments, as shown in
Table~\ref{tab:param}. We varied (1) the number of workers to assess the
scalability of the engines, (2) the number of BigBrain blocks in
Incrementation and Histogram to measure the effect of different IO patterns
and parallelization degrees, and (3) the number of iterations and sleep
delay in Incrementation to evaluate the effect of job length.

>>>>

We used three neuroimaging applications to evaluate the engines in different
conditions. The first two ones, incrementation and histogram, are simple synthetic
benchmarks representing basic map-only and map-reduce applications. The third one is
a realistic application representative of popular BIDS
applications~\cite{gorgolewski2017bids}.

%% Incrementation
\subsubsection{Incrementation}
We used an adaptation of the simple image incrementation pipeline used
in~\cite{hayot2019performance} (see Algorithm~\ref{alg:incrementation}).
The application reads blocks of the BigBrain image from the shared file
system, increments the intensity value of each voxel by 1 to avoid caching
effects, sleeps for a configurable amount of time to emulate a more complex
processing, repeats this process for a specified amount of iterations, and
finally writes the result as a NifTI image back to the shared file system.
This application allows us to study the behavior of the engines when all
inputs are processed independently, in a map-only way (see
Figure~\ref{fig:tg-inc}). This mimics the behavior of analyzing multiple
independent subjects in parallel.

@@

\begin{figure}[!b]
    \centering
    \includegraphics[width=0.125\textwidth,
    angle=-90]{images/incrementation-task-graph.png}
    \caption{Task graph for Incrementation with 5 iterations and 3 BigBrain blocks.
    Circles represent the incrementation and sleep function while rectangles
    represent stages of the image blocks.}\label{fig:tg-inc}
\end{figure}

@@

 Our second application calculates the histogram of the BigBrain image (see
 Algorithm~\ref{alg:histogram}). It reads the image blocks from the shared file
 system, calculates intensity frequencies, aggregates the frequencies across the
 blocks, and finally writes the resulting histogram to the shared file system as a
 single \SI{766}{\kilo\byte} file. This application implements a typical map-reduce
 pattern, where the final result is obtained from all the input blocks (see
 Fig.~\ref{fig:tg-histo}). This application requires data shuffling thus inter-worker
 communication. The total amount of shuffled data is however limited to
 \SI{2.62}{\mega\byte} per blocks as it only consists of image histograms.

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.16\textwidth, angle=-90]{images/histogram-task-graph.png}
    \caption{Task graph for Histogram with 3 BigBrain blocks. Circles represent the
    functions while rectangles represent stages of the image blocks.
    }\label{fig:tg-histo}
\end{figure}

@@

We used the \href{https://github.com/BIDS-Apps/example}{BIDS app example}
application available on GitHub. This application operates in a map-reduce
way. The map phase, also called participant analysis, extracts the brain
tissues of a subject's 3D MRI using the popular FMRIB Software Library
(\href{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki}{FSL}), and writes the
resulting image (\TG{X}MB on average) to the shared file system. The reduce
phase, also called group analysis, computes the volume of each brain and
returns the average volume, shuffling a total of \TG{X}GB image data.
\MD{Determine values when dataset is determined for the application.}

@@

We varied four parameters in our experiments, as shown in
Table~\ref{tab:param}. We varied (1) the number of workers to assess the
scalability of the engines, (2) the number of BigBrain blocks in
Incrementation and Histogram to measure the effect of different IO patterns
and parallelization degrees, and (3) the number of iterations and sleep
delay in Incrementation to evaluate the effect of job length.
We run the experiments in a non-deterministic order and clear the cache of each
worker before each benchmark.
